Spike Spectra for Recurrences

### 2022-11-16 Manuscript/entropy-1919488  proof/entropy-1919488.tex
changed figures

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, 14473 Potsdam, Germany; \hl{hellmann@pik-potsdam.de (F.H.); anvari@pik-potsdam.de (M.A.); kurths@pik-potsdam.de (J.K.)} %MDPI: We added the email addresses here according to the submitting system. Please confirm.%Authors: we agree
\\
$^{2}$ \quad Institute of Physics and Astronomy, University of Potsdam, 14476 Potsdam, Germany\\
\hl{$^{3}$} %MDPI: The affiliation numbers should appear in numerical order. Please check the suggested changes.%Authors: we agree
 \quad Institute of Physics, Humboldt Universit\"at zu Berlin, 12489 Berlin, Germany\\
\hl{$^{4}$} \quad Institute of Geosciences, University of Potsdam, 14476 Potsdam, Germany}

@@

\citet{Zbilut2008} found that $\tau$-RR could be interpreted as analogous to the auto-correlation function $C(\tau)$ and, hence, via the Wiener--Khinchin theorem, provide an analogue,
``generalized'', spectral density. This is reasonable, since the average distances for a given lag $\tau$ 
\begin{equation}
\overline{D}(\tau) = \frac{1}{N-\tau}\sum_{i=1}^{N-\tau} D_{i, i+\tau}
\end{equation}
can be directly read from the distance matrix $\mathbf{D}$ and are also preserved in its thresholded version $\tau$-RR. There are clear advantages to recurrence-derived spectral density, 
i.e., Fourier transforming (FT) $\tau$-RR (Figure~\ref{fig_tau_rr_spectrum_example}D,E), instead 
of $C(\tau)$. There are no assumptions for stationarity or sampling when constructing a RP.
Furthermore, since an RP can represent high-dimensional dynamics and its $\tau$-RR serves as a plug-in for $C(\tau)$, the correlation structures of higher dimensional spaces can be read from 
the recurrence-derived Fourier~spectrum.\\
\vspace{-6pt}
\begin{figure}[H]
 %\centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_spectrum_example}
 \caption{\hl{Schematic} %MDPI: Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 illustration of a $\tau$-recurrence-rate-based spectrum. (\textbf{A}) $x$-component time series of the Lorenz63-System (Equation~\eqref{eq_model_Lorenz63}) and 
 (\textbf{B}) its corresponding Fourier power spectrum. 
 (\textbf{C})~Reconstructed state space portrait from the time series shown in (\textbf{A}) using PECUZAL time-delay \hl{embedding} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10.
 \cite{Kraemer2021}. 
 (\textbf{D}) Subset of the recurrence plot and the corresponding $\tau$-recurrence rate obtained from the state space trajectory in (\textbf{C}). 
(\textbf{E}) Fourier Power spectrum obtained from the $\tau$-recurrence rate (subset shown in panel (\textbf{D}\hl{))} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10.
 \cite{Zbilut2008}. (\textbf{D},\textbf{E}) show the results of a part of the time series, 
 which is highlighted in pink in (\textbf{A}).
 }\label{fig_tau_rr_spectrum_example}
\end{figure}

@@

For example, consider the signal we would like to analyze 
(e.g., the $\tau$-RR of a system) to be a Dirac comb (DC) with inter-spike period $T_\text{is}$: 
\begin{equation}
\text{DC}_{\text{is}}(t) = \sum_{k=-\infty}^{\infty} \delta(t-kT_\text{is}),
\label{eq_dirac_comb}
\end{equation}
i.e., a series of Dirac delta functions for a period $T_\text{is}$. There is only one single period, $T_\text{is}$, in this signal (Figure~\ref{fig_tau_rr_dirac_comb}A,D); 
in principle, we would strive for a single peak in the frequency domain of this signal at frequency $f=1/T_\text{is}$. The Fourier spectrum does not meet this expectation 
and instead of a single frequency, there are many excited frequencies (Figure~\ref{fig_tau_rr_dirac_comb}B,E). This is because the Fourier components constructively contribute
to every frequency $1/T_\text{is}$; therefore, $\text{DC}_{\text{is}}(t)$ coincides with its own Fourier transform up to a factor $1/T_\text{is}$ \cite{Norden1998}. ({\hl{This can also } %MDPI: footnote is not allowed in this journal, we moved here, please confirm
be observed for neuron spike trains, e.g., \cite{Orcioni2020,Biagetti2017}.})

@@

\begin{figure}[H]
% \centering
 \includegraphics[width=0.76\textwidth]{./figures/fig_tau_rr_example_roessler}
 \caption{\hl{Inter-spike spectra} %MDPI: Please use a consistent numbering style for the subfigures throughout the paper. P is after N;  Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1".
 of the $\tau$-RR of the R\"ossler system in three different dynamical regimes with parameters $b=2$, $c=4$. 
 (\textbf{A}) Trajectory of the system in a period-2 (parameter $a=0.36$), (\textbf{B})~in a period-3 (parameter $a=0.41$) and 
 (\textbf{C}) in a chaotic regime (parameter $a=0.428$). 
\mbox{ (\textbf{D}--\textbf{F})}~The corresponding RPs, obtained by using a recurrence threshold corresponding to a 10\% global 
 recurrence rate for (\textbf{D},\textbf{E}) and 5\% for (\textbf{F}). 
  (\textbf{G}--\textbf{I}) $\tau$-RRs of the shown RPs. 
  (\textbf{J}--\textbf{L}) The proposed inter-spike spectra of the $\tau$-RRs shown in panels (\textbf{G}--\textbf{I}). Spectra were obtained with a LASSO regression and a regularization threshold 
  corresponding to $\rho=0.95$ accordance of $\tau$-RRs and re-composed signals. The distance ratio of the peaks reflect the limit cycle dynamic.  
  (\textbf{M},\textbf{N},\textbf{P}) Fourier power spectra of the $\tau$-RRs shown in panels (\textbf{G}--\textbf{I}). }
\label{fig_tau_rr_example_roessler}
\end{figure}

@@

\hl{In this example} %MDPI: we add indent here, please confirm
, the null hypothesis for constructing the surrogate data is that the data stem from a process which yields the same auto-correlation, and
hence, the same Fourier power spectral density, and the same 
amplitude distribution. We consider the number of significant peaks in the inter-spike spectrum with
respect to the control parameter in order to distinguish the corresponding dynamics (Figure~\ref{fig_tau_rr_logistic}C).
A correlation with the positive Lyapunov exponent (Figure~\ref{fig_tau_rr_logistic}A) is discernible 
($\rho_{\text{Pearson}}(\text{Lyapunov})=0.72$). This analysis can handle period-doubling since it measures the dominant cycles via the inter-spike spectrum. 
However, whenever the periods of the new cycles coincide with integer multiples of the periods of already existing cycles, the approach cannot detect 
period doubling. A similar case can be observed in Figure~\ref{fig_tau_rr_example_roessler}J,K, where the number of peaks does not change, but  the mutual distance does.

@@

\begin{figure}[H]
 %\centering
 \includegraphics[width=\textwidth]{./figures/fig_tau_rr_logistic}
 \caption{(\textbf{A}) Bifurcation diagram and Lyapunov exponent and of the Logistic map as a function of the control parameter $r$.
 (\textbf{B}) Number of significant peaks ($\alpha=0.05$) in the inter-spike spectrum of the $\tau$-RR and its Pearson correlation coefficient to the Lyapunov exponent shown in (\textbf{A})
 (white noise surrogates). 
 (\textbf{C}) Same as (\textbf{B}), but for iterative Amplitude Adjusted Fourier Transform (iAAFT) \hl{surrogates} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10.
 \cite{Schreiber1996,Schreiber2000}. To obtain the inter-spike spectra, we used 
 a LASSO regression and a regularization threshold corresponding to the $\rho=0.95$ accordance of $\tau$-RRs and re-composed signals.  
}
\label{fig_tau_rr_logistic}
\end{figure}

@@

\begin{figure}[H]
% \centering
 \includegraphics[width=\textwidth]{./figures/fig_power_grid_time_series}
 \caption{Subset of the frequency time series of (\textbf{A}) Central Europe \hl{(CE)} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10.
 \cite{haehne2018footprint} and (\textbf{C}) Great Britain \hl{(GB)} %MDPI: 
 \cite{GB} along with their autocorrelation functions in (\textbf{B},\textbf{D}), respectively. The normalized autocorrelations in (\textbf{B},\textbf{D}) were computed with the entire time series, whereas the subsets shown in (\textbf{A},\textbf{C}) 
 contain only $1441$ samples.}
\label{fig_power_grid_time_series}
\end{figure}

@@

 \includegraphics[width=0.93\textwidth]{./figures/A3B}
 \caption{\hl{Same as in} %MDPI: please add explanation for A--I
 Figure~\ref{fig_tau_rr_example_roessler}, but here with 5\% additive Gaussian white noise on each component $x$, $y$ and $z$. The appearance of the inter-spike spectra 
 in (\textbf{J}--\textbf{L}), and the Fourier spectra in (\textbf{M},\textbf{N},\textbf{P}) are unaffected by the additive noise.}
\label{fig_tau_rr_example_roessler_noise}
\end{figure}

@@

\bibitem[Zhang {et~al.}(2022)Zhang, Widmann, and Barreira]{Zhang2022}
\textls[-15]{Zhang, S.; Widmann, D.; Barreira, D.S}.
\newblock {JuliaOptimalTransport/OptimalTransport.jl: v0.3.19}%MDPI: Please provide more information about the type of the article, such as book (please provide the name of the publisher and location of it), online resource (please provide the URL of the website and the date it was accessed (day, month, year)), or journal article (please provide the name of the journal, the year and volume it published, and the page number).
. {2022}.
\newblock
   \hl{Available online:} %MDPI: we replace the correct web link, please confirm
 \url{https://zenodo.org/record/5920148\#.Y3SbbORByUk} \hl{(accessed on)} %MDPI: Please add accessed date.

>>>>

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, 14473 Potsdam, Germany; \hl{hellmann@pik-potsdam.de (F.H.); anvari@pik-potsdam.de (M.A.); kurths@pik-potsdam.de (J.K.)} %MDPI: We added the email addresses here according to the submitting system. Please confirm.%Authors: we agree
\\
$^{2}$ \quad Institute of Physics and Astronomy, University of Potsdam, 14476 Potsdam, Germany\\
\hl{$^{3}$} %MDPI: The affiliation numbers should appear in numerical order. Please check the suggested changes. %Authors: we agree
 \quad Institute of Physics, Humboldt Universit\"at zu Berlin, 12489 Berlin, Germany\\
\hl{$^{4}$} \quad Institute of Geosciences, University of Potsdam, 14476 Potsdam, Germany}

@@

\citet{Zbilut2008} found that $\tau$-RR could be interpreted as analogous to the auto-correlation function $C(\tau)$ and, hence, via the Wiener--Khinchin theorem, provide an analogue,
``generalized'', spectral density. This is reasonable, since the average distances for a given lag $\tau$ 
\begin{equation}
\overline{D}(\tau) = \frac{1}{N-\tau}\sum_{i=1}^{N-\tau} D_{i, i+\tau}
\end{equation}
can be directly read from the distance matrix $\mathbf{D}$ and are also preserved in its thresholded version $\tau$-RR. There are clear advantages to recurrence-derived spectral density, 
i.e., Fourier transforming (FT) $\tau$-RR (Figure~\ref{fig_tau_rr_spectrum_example}D,E), instead 
of $C(\tau)$. There are no assumptions for stationarity or sampling when constructing a RP.
Furthermore, since an RP can represent high-dimensional dynamics and its $\tau$-RR serves as a plug-in for $C(\tau)$, the correlation structures of higher dimensional spaces can be read from 
the recurrence-derived Fourier~spectrum.\\
\vspace{-6pt}
\begin{figure}[H]
 %\centering
 \includegraphics[width=0.9\textwidth]{./figures/fig_tau_rr_spectrum_example}
 \caption{\hl{Schematic} %MDPI: Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1". %Authors: we changed to minus sign
 illustration of a $\tau$-recurrence-rate-based spectrum. (\textbf{A}) $x$-component time series of the Lorenz63-System (Equation~\eqref{eq_model_Lorenz63}) and 
 (\textbf{B}) its corresponding Fourier power spectrum. 
 (\textbf{C})~Reconstructed state space portrait from the time series shown in (\textbf{A}) using PECUZAL time-delay \hl{embedding} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10. %Authors: we confirm that we hold the copyright of this figure
 \cite{Kraemer2021}. 
 (\textbf{D}) Subset of the recurrence plot and the corresponding $\tau$-recurrence rate obtained from the state space trajectory in (\textbf{C}). 
(\textbf{E}) Fourier Power spectrum obtained from the $\tau$-recurrence rate (subset shown in panel (\textbf{D}\hl{))} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10. %Authors: we confirm that we hold the copyright of this figure
 \cite{Zbilut2008}. (\textbf{D},\textbf{E}) show the results of a part of the time series, 
 which is highlighted in pink in (\textbf{A}).
 }\label{fig_tau_rr_spectrum_example}
\end{figure}

@@

For example, consider the signal we would like to analyze 
(e.g., the $\tau$-RR of a system) to be a Dirac comb (DC) with inter-spike period $T_\text{is}$: 
\begin{equation}
\text{DC}_{\text{is}}(t) = \sum_{k=-\infty}^{\infty} \delta(t-kT_\text{is}),
\label{eq_dirac_comb}
\end{equation}
i.e., a series of Dirac delta functions for a period $T_\text{is}$. There is only one single period, $T_\text{is}$, in this signal (Figure~\ref{fig_tau_rr_dirac_comb}A,D); 
in principle, we would strive for a single peak in the frequency domain of this signal at frequency $f=1/T_\text{is}$. The Fourier spectrum does not meet this expectation 
and instead of a single frequency, there are many excited frequencies (Figure~\ref{fig_tau_rr_dirac_comb}B,E). This is because the Fourier components constructively contribute
to every frequency $1/T_\text{is}$; therefore, $\text{DC}_{\text{is}}(t)$ coincides with its own Fourier transform up to a factor $1/T_\text{is}$ \cite{Norden1998}. ({\hl{This can also } %MDPI: footnote is not allowed in this journal, we moved here, please confirm  %Authors: we agree/ confirm
be observed for neuron spike trains, e.g., \cite{Orcioni2020,Biagetti2017}.})

@@

\begin{figure}[H]
% \centering
 \includegraphics[width=0.76\textwidth]{./figures/fig_tau_rr_example_roessler}
 \caption{\hl{Inter-spike spectra} %MDPI: Please use a consistent numbering style for the subfigures throughout the paper. P is after N;  Please change the hyphen (-) into minus sign ($-$, "U+2212"). e.g., "-1" should be "$-$1". %Authors: We changed the hyphen to minus sign. We have modified the panel numbering to make it consistent with the other figures.
 of the $\tau$-RR of the R\"ossler system in three different dynamical regimes with parameters $b=2$, $c=4$. 
 (\textbf{A}) Trajectory of the system in a period-2 (parameter $a=0.36$), (\textbf{B})~in a period-3 (parameter $a=0.41$) and 
 (\textbf{C}) in a chaotic regime (parameter $a=0.428$). 
\mbox{ (\textbf{D}--\textbf{F})}~The corresponding RPs, obtained by using a recurrence threshold corresponding to a 10\% global 
 recurrence rate for (\textbf{D},\textbf{E}) and 5\% for (\textbf{F}). 
  (\textbf{G}--\textbf{I}) $\tau$-RRs of the shown RPs. 
  (\textbf{J}--\textbf{L}) The proposed inter-spike spectra of the $\tau$-RRs shown in panels (\textbf{G}--\textbf{I}). Spectra were obtained with a LASSO regression and a regularization threshold 
  corresponding to $\rho=0.95$ accordance of $\tau$-RRs and re-composed signals. The distance ratio of the peaks reflect the limit cycle dynamic.  
  (\textbf{M},\textbf{N},\textbf{P}) Fourier power spectra of the $\tau$-RRs shown in panels (\textbf{G}--\textbf{I}). }
\label{fig_tau_rr_example_roessler}
\end{figure}

@@

\hl{In this example} %MDPI: we add indent here, please confirm %Authors: we agree/ confirm
, the null hypothesis for constructing the surrogate data is that the data stem from a process which yields the same auto-correlation, and
hence, the same Fourier power spectral density, and the same 
amplitude distribution. We consider the number of significant peaks in the inter-spike spectrum with
respect to the control parameter in order to distinguish the corresponding dynamics (Figure~\ref{fig_tau_rr_logistic}C).
A correlation with the positive Lyapunov exponent (Figure~\ref{fig_tau_rr_logistic}A) is discernible 
($\rho_{\text{Pearson}}(\text{Lyapunov})=0.72$). This analysis can handle period-doubling since it measures the dominant cycles via the inter-spike spectrum. 
However, whenever the periods of the new cycles coincide with integer multiples of the periods of already existing cycles, the approach cannot detect 
period doubling. A similar case can be observed in Figure~\ref{fig_tau_rr_example_roessler}J,K, where the number of peaks does not change, but  the mutual distance does.

@@

\begin{figure}[H]
 %\centering
 \includegraphics[width=\textwidth]{./figures/fig_tau_rr_logistic}
 \caption{(\textbf{A}) Bifurcation diagram and Lyapunov exponent and of the Logistic map as a function of the control parameter $r$.
 (\textbf{B}) Number of significant peaks ($\alpha=0.05$) in the inter-spike spectrum of the $\tau$-RR and its Pearson correlation coefficient to the Lyapunov exponent shown in (\textbf{A})
 (white noise surrogates). 
 (\textbf{C}) Same as (\textbf{B}), but for iterative Amplitude Adjusted Fourier Transform (iAAFT) \hl{surrogates} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10. %Authors: we confirm that we hold the copyright of this figure
 \cite{Schreiber1996,Schreiber2000}. To obtain the inter-spike spectra, we used 
 a LASSO regression and a regularization threshold corresponding to the $\rho=0.95$ accordance of $\tau$-RRs and re-composed signals.  
}
\label{fig_tau_rr_logistic}
\end{figure}

@@

\begin{figure}[H]
% \centering
 \includegraphics[width=\textwidth]{./figures/fig_power_grid_time_series}
 \caption{Subset of the frequency time series of (\textbf{A}) Central Europe \hl{(CE)} %MDPI: Please ensure that permission has been obtained and there is no copyright issue. If copyright is needed, please provide a citation in the following format: "Reprinted/adapted with permission from Ref. [XX]. Copyright year, copyright owner's name". More details on "Copyright and Licensing" are available via the following link: https://www.mdpi.com/ethics#10. %Authors: we confirm that we hold the copyright of this figure
 \cite{haehne2018footprint} and (\textbf{C}) Great Britain \hl{(GB)} %MDPI: 
 \cite{GB} along with their autocorrelation functions in (\textbf{B},\textbf{D}), respectively. The normalized autocorrelations in (\textbf{B},\textbf{D}) were computed with the entire time series, whereas the subsets shown in (\textbf{A},\textbf{C}) 
 contain only $1441$ samples.}
\label{fig_power_grid_time_series}
\end{figure}

@@

 \includegraphics[width=0.93\textwidth]{./figures/A3B}
 \caption{\hl{Same as in} %MDPI: please add explanation for A--I
 Figure~\ref{fig_tau_rr_example_roessler}, but here with 5\% additive Gaussian white noise on each component $x$, $y$ and $z$. 
 (\textbf{A}) Trajectory of the system in a period-2 (parameter $a=0.36$), (\textbf{B})~in a period-3 (parameter $a=0.41$) and 
 (\textbf{C}) in a chaotic regime (parameter $a=0.428$). 
\mbox{ (\textbf{D}--\textbf{F})}~The corresponding RPs, obtained by using a recurrence threshold corresponding to a 10\% global 
 recurrence rate for (\textbf{D},\textbf{E}) and 5\% for (\textbf{F}). 
  (\textbf{G}--\textbf{I}) $\tau$-RRs of the shown RPs, 
  (\textbf{J}--\textbf{L}) the corresponding inter-spike spectra, and
  (\textbf{M},\textbf{N},\textbf{P}) the Fourier power spectra. 
 The appearance of the inter-spike spectra 
 in (\textbf{J}--\textbf{L}), and the Fourier spectra in (\textbf{M},\textbf{N},\textbf{P}) are unaffected by the additive noise.}
\label{fig_tau_rr_example_roessler_noise}
\end{figure}

@@

\bibitem[Zhang {et~al.}(2022)Zhang, Widmann, and Barreira]{Zhang2022}
\textls[-15]{Zhang, S.; Widmann, D.; Barreira, D.S}.
\newblock {JuliaOptimalTransport/OptimalTransport.jl: v0.3.19}%MDPI: Please provide more information about the type of the article, such as book (please provide the name of the publisher and location of it), online resource (please provide the URL of the website and the date it was accessed (day, month, year)), or journal article (please provide the name of the journal, the year and volume it published, and the page number). %Authors: we confirm that we hold the copyright of this figure
. {2022}.
\newblock
   \hl{Available online:} %MDPI: we replace the correct web link, please confirm
 \url{https://zenodo.org/record/5920148\#.Y3SbbORByUk} \hl{(accessed on)} %MDPI: Please add accessed date.
